---
title: "Text Mining Analysis"
output:
  html_document:
    df_print: paged
date: "2025-05-28"
---

# Text Mining Analysis

This document provides a general workflow for analysing a corpus of text documents using text mining, hierarchical clustering, sentiment analysis and network analysis.

The analysis is designed to be **customisable**. Users can replace the example corpus with their own collection of `.txt` documents and modify the settings below depending on the size and characteristics of their corpus.

Throughout the analysis, the user should interpret the results in the context of their own documents rather than assuming that the patterns identified in an example corpus will apply to every corpus.

---

# Document-Term Matrix

## Creating and preprocessing the corpus

A Document-Term Matrix (DTM) represents the relationship between documents and the terms contained within them.

Each **row** represents a document and each **column** represents a token/term. The values indicate how frequently each term occurs in each document.

Before constructing the DTM, the documents are transformed to reduce unnecessary variation caused by formatting, punctuation, capitalisation and common words.

The exact preprocessing required depends on the corpus. For example, hyphenated terms may need to be transformed if the hyphen prevents important words from being recognised correctly.

```{r setup, message=FALSE, warning=FALSE}
# Clear objects from the current R environment.
rm(list = ls())

# Install packages if they are not already installed.
install.packages("tm")
install.packages("slam")
install.packages("SnowballC")
install.packages("proxy")
install.packages("SentimentAnalysis")
install.packages("igraph")

# Load packages required for the analysis.
library(slam)
library(tm)
library(SnowballC)
library(proxy)
library(SentimentAnalysis)
library(igraph)
```

## Customisation: selecting the corpus

The corpus should be stored in a folder containing the `.txt` documents to be analysed. For example:
CorpusAbstracts/
Book1.txt
Book2.txt
Sports1.txt
Sports2.txt

```{r}
# Create corpus
cname <- file.path(".", "CorpusAbstracts")

if (!dir.exists(cname)) {
  stop("Cannot find CorpusAbstracts folder.")
}

# Summary of docs to analyse
docs <- Corpus(DirSource(cname))
summary(docs)
```

The following transformations are applied:

1. Hyphens are replaced with spaces.
2. Numbers are removed.
3. Punctuation is removed.
4. All words are converted to lowercase.
5. English stop words are removed.
6. Extra whitespace is removed.
7. Words are stemmed.

These transformations reduce differences between words that are not useful for identifying relationships between documents.

For example, converting words to lowercase ensures that `"Travel"` and `"travel"` are treated as the same token.

Stemming reduces related forms of words to a common root. For example, words such as `"connect"`, `"connected"` and `"connecting"` may be reduced to a common stem.

The appropriate preprocessing depends on the corpus. If a particular punctuation mark, number, abbreviation or formatting feature contains meaningful information, removing it may not be appropriate.

```{r}
# Specific text transformations
# Hyphen to space
toSpace = content_transformer(function(x, pattern)
gsub(pattern, " ", x))
docs = tm_map(docs, toSpace, "-")
# Tokenisation and stemming 
docs = tm_map(docs, removeNumbers)
docs = tm_map(docs, removePunctuation)
docs = tm_map(docs, content_transformer(tolower))
docs = tm_map(docs, removeWords, stopwords("english"))
docs = tm_map(docs, stripWhitespace)
docs = tm_map(docs, stemDocument, language =
"english")
```

## Creating the Document-Term Matrix

The DTM is created from the processed corpus. Initially, the DTM may contain a large number of terms, including terms that occur in only a small number of documents.These sparse terms can therefore be removed.

The sparsity threshold determines how much of the DTM must contain zero values before a term is removed. The threshold should generally be adjusted through trial and error to obtain a manageable number of useful terms.

```{r}
# Create document term matrix
dtm <- DocumentTermMatrix(docs)
inspect(dtm)

# The first value represents the number of documents and the second represents the number of terms.
dim(dtm)

# Remove sparse terms i.e Cols with at least 45% empty cells
dtms <- removeSparseTerms(dtm, 0.45)
inspect(dtms)

# Display the reduced DTM as a table.
dtmsx <- as.matrix(dtms)
knitr::kable(dtmsx)
```

### Interpreting the DTM

The resulting table can be used to determine which terms occur in which documents. A large value in a cell indicates that the corresponding term occurs frequently in that document. A zero indicates that the term does not occur in that document.

The DTM therefore provides the basis for the later analyses. The same document-term relationships are subsequently used to calculate document similarity, construct networks and investigate relationships between documents and tokens.

---

# Hierarchical Clustering

## Cosine distance

Hierarchical clustering is used to identify groups of documents based on similarities in their term usage. Cosine distance is used to measure how different two documents are. Each document can be represented as a vector of term frequencies. Cosine similarity focuses on the **direction of these vectors**, rather than simply their magnitude.

This is useful for text analysis because two documents can discuss similar topics while having very different lengths. Cosine distance is therefore based on the pattern of terms used by documents rather than simply the total number of words they contain. A smaller cosine distance indicates that two documents have more similar term patterns, while a larger distance indicates greater dissimilarity.

Hierarchical clustering uses the calculated distances to progressively combine similar documents into groups. The resulting dendrogram provides a visual representation of these relationships.

```{r fig.height=15, fig.width=20}
# Change k according to the number of clusters you want to investigate
k <- 2

# Calculate pairwise cosine distances between documents
distmatrix <- proxy::dist(dtmsx, method = 'cosine')
distmatrix

# Ward's method attempts to form clusters by minimising within-cluster variation
fit = hclust(distmatrix, method = 'ward.D')
# Plot the dendrogram
plot(fit, hang = -1, main = 'Abstracts Cosine Distance')

# Create clusters
groups <- cutree(fit, k)
# Display the cluster assigned to each document
groups

# The following assumes filenames contain a category followed by a number, for example:
# book1.txt
# sports2.txt
# travel5.txt
#
# Modify the pattern if your filenames use another structure.
doc_names <- rownames(dtmsx)
groups_known <- sub("[0-9]+\\.txt$", "", doc_names)

# Inspect the document/category mapping.
data.frame(
  Document = doc_names,
  KnownGroup = groups_known,
  Cluster = groups)
```

## Calculating cluster accuracy with known categories

If the corpus already has known categories, these can be compared with the clusters produced by the algorithm.

* **Cosine distance** determines how similar the documents are.
* **Hierarchical clustering** uses those distances to construct clusters.
* **Known categories/genres** can then be used to evaluate how closely the resulting clusters correspond to the expected categories.

If each cluster is assigned to the category that appears most frequently within that cluster, an approximate classification accuracy can be calculated. The calculation should be treated as an evaluation of how closely the automatically generated clusters correspond to the known categories. The accuracy value indicates the proportion of documents whose automatically assigned cluster corresponds to the category assigned using the known labels. Accuracy from a confusion matrix is calculated by dividing the total number of correct predictions by the total number of all predictions.

The dendrogram and confusion matrix should be considered alongside the actual content of the documents. Users can investigate:

* which documents appear together;
* which documents are separated from the others;
* which terms are shared by documents in the same cluster;
* whether documents in the same cluster discuss similar subjects;
* whether documents from the same known category have similar writing patterns;
* whether documents from different categories have overlapping vocabulary.

For example, if several documents from the same category appear in one cluster, this may indicate that they share vocabulary or subject matter. Clusters can also be driven by a common writing style, named entity, event, phrase or other vocabulary pattern rather than the category itself. The qualitative interpretation should therefore explain **why particular documents appear similar**, using examples from the corpus where appropriate.

```{r}
# Create a confusion matrix showing the relationship between known document categories and clustering assignments
TA <- table(
  GroupNames = groups_known,
  Clusters = groups)

TA

# Calculate approximate clustering accuracy by taking the largest group within each cluster as the correctly classified documents
cluster_accuracy <- function(TA) {
  correct <- sum(apply(TA, 2, max))
  total <- sum(TA)
  accuracy <- correct / total
  return(accuracy)
}

cluster_accuracy(TA)
```
---

# Sentiment Analysis

Sentiment analysis attempts to quantify the emotional or evaluative characteristics of each document. The `SentimentAnalysis` package provides several dictionary-based measures. A dictionary method compares words in each document with words contained in a predefined sentiment dictionary. Different dictionaries are constructed differently and therefore measure slightly different aspects of sentiment.

For this analysis, multiple sentiment measures can initially be inspected before selecting the measure most relevant to the research question. The following plots compare sentiment-related measures across the known document groups.

```{r fig.height=17, fig.width=20}
# Analyse sentiment using the processed corpus
sentiment_docs <- analyzeSentiment(docs)

# Create data frame from corpus summary
names <- as.data.frame(as.table(summary(docs)))
names$Topic <- sub("[0-9]+\\.txt$", "", names$Var1)
names <- names[, c("Var1", "Topic")]
colnames(names) <- c("Document", "Group")
sentiment_docs <- cbind(names, sentiment_docs)

sentiment_docs
```


```{r fig.height=17, fig.width=20}
# Plot sentiment measures
par(mfrow= c(2,2))
# Compare word counts between groups
boxplot(WordCount ~ Group, data = sentiment_docs, frame = TRUE, main = "Word Count by Group")
# Compare QDAP sentiment scores
boxplot(SentimentQDAP ~ Group, data = sentiment_docs, frame = TRUE, main = "Sentiment (QDAP) by Group")
# Compare QDAP positivity
boxplot(PositivityQDAP ~ Group, data = sentiment_docs, frame = TRUE, main = "Positivity (QDAP) by Group")
# Compare uncertainty using the LM measure
boxplot(RatioUncertaintyLM ~ Group, data = sentiment_docs, frame = TRUE, main = "Uncertainty Ratio (LM) by Group")
```
### Interpreting sentiment measures

The exact interpretation depends on the dictionary and measure being used. For example:

* **WordCount** describes document length
* **PositivityQDAP** represents the positivity identified by the QDAP dictionary
* **SentimentQDAP** provides a sentiment-related score based on QDAP
* **RatioUncertaintyLM** represents the proportion of words identified as expressing uncertainty under the LM dictionary

The boxplots allow the user to compare both the **central tendency** and **variability** of the measures across groups. A higher median indicates a generally higher observed score, while a wider box or greater range indicates greater variation between documents. The boxplots should be examined before conducting statistical tests because they provide a visual indication of potential differences and unusual observations.

## Statistical testing

Statistical tests can be used to investigate specific hypotheses suggested by the research question and exploratory plots. For example, a researcher may wish to investigate whether one group's average positivity is greater than another group's average positivity.

A one-sided t-test evaluates whether the observed difference is consistent with a specified directional hypothesis. The resulting **p-value** represents how unusual the observed data would be under the null hypothesis. A small p-value provides evidence against the null hypothesis in favour of the specified alternative hypothesis. The chosen significance level should be stated before interpreting the result.

Importantly, statistical significance does not by itself indicate the size or practical importance of a difference. The group means and the magnitude of the difference should also be examined.

An F-test can be used to compare the variances of two groups. For a two-sided F-test:

* The null hypothesis states that the two population variances are equal.
* The alternative hypothesis states that the population variances are different.

If the p-value is small, the data provide evidence against equal variances. If the p-value is not small, there is insufficient evidence to conclude that the variances differ.


```{r}
# Example:
# Is the average positivity score for Group A greater than the average positivity score for Group B?
group_A <- sentiment_docs[sentiment_docs$book == "GROUP_A", ]
group_B <- sentiment_docs[sentiment_docs$sport == "GROUP_B", ]

t.test(
  group_A$PositivityQDAP,
  group_B$PositivityQDAP,
  alternative = "greater"
)

# Example:
# Is the average negativity score for Group A less than the average negativity score for Group B?
t.test(
  group_A$NegativityQDAP,
  group_B$NegativityQDAP,
  alternative = "less"
)

# Compare variability between two groups
var.test(
  group_A$RatioUncertaintyLM,
  group_B$RatioUncertaintyLM
)
```
---

# Single-Mode Document Network

A single-mode document network represents **documents as nodes**. An edge is created between two documents when they share terms. The weight of an edge represents the number of shared terms.

Therefore:
* nodes = documents;
* edges = relationships between documents;
* edge weight = number of shared terms.

Clustering asks which documents can be grouped based on their distances, whereas network analysis allows relationships between individual documents to be visualised and investigated.

```{r fig.height=7, fig.width=20}
# Convert to binary matrix
# A value of 1 means that a term occurs in a document, while 0 means that it does not
dtmsb <- as.matrix((dtmsx > 0) + 0)
# Multiply binary matrix by its transpose, this calculates how many terms each pair of documents has in common
absmatrix = dtmsb %*% t(dtmsb)
# Make leading diagonal zero, a document should not be connected to itself
diag(absmatrix) = 0

knitr::kable(absmatrix)
```

## Network statistics

Several network measures can be calculated to describe the structure of the document network.

**Degree**
Degree is the number of edges connected to a node. A document with high degree shares terms with many other documents. If the graph is completely connected, every document is connected to every other document and therefore degree will equal `n - 1`.

**Betweenness centrality**
Betweenness measures how often a node lies along shortest paths between other nodes. A document with high betweenness may act as a bridge between different parts of the network. This does not necessarily mean that the document is the most similar to all other documents.

**Closeness centrality**
Closeness measures how near a node is to other nodes through the network. A higher value indicates that a node can generally reach other nodes through relatively short paths.

**Eigenvector centrality**
Eigenvector centrality considers not only how many connections a node has, but also how important the nodes it connects to are. A document can therefore have high eigenvector centrality when it is connected to other highly connected/influential documents.

**Diameter**
The diameter is the longest shortest path between any two nodes. It describes the maximum separation between nodes in the network.

**Average path length**
This is the average shortest-path distance between nodes. Lower values indicate that nodes are generally closer together.

**Density**
Network density represents the proportion of possible edges that actually exist. A density of 1 means that every possible pair of nodes is connected.

**Transitivity**
Transitivity measures the tendency for connected nodes to form triangles. A high value indicates a strong tendency for nodes with shared connections to also be connected to one another.

```{r}
# Create undirected weighted graph
abs_single = graph_from_adjacency_matrix(absmatrix, mode = "undirected", weighted = TRUE)
plot(abs_single)

# Networkk summary
cat('Is simple?',is_simple(abs_single))
cat('\nIs connected?',is_connected(abs_single))
cat('\nDiameter?', diameter(abs_single))
cat('\nAvg path length?',mean_distance(abs_single))
cat('\nEdge density?',edge_density(abs_single))
cat('\nTransitivity?',transitivity(abs_single))

# Centrality measures
bet_abs = betweenness(abs_single)
close_abs = closeness(abs_single)
deg_abs = degree(abs_single)
eig_abs = eigen_centrality(abs_single)$vector
abs_table = cbind(deg_abs, bet_abs, close_abs, eig_abs)
knitr::kable(abs_table)
```
## Improving the network visualisation

The basic graph can be improved by encoding additional information visually. For example:
* node colour = hierarchical cluster;
* node size = eigenvector centrality;
* edge width = shared-term weight;
* layout = Fruchterman-Reingold.

The improved graph below allows several features to be examined simultaneously. Documents that appear close together may have similar network relationships, although physical distance in a force-directed layout should not automatically be interpreted as a statistical distance. Larger nodes indicate greater eigenvector centrality under the selected visualisation. Thicker edges indicate more shared terms. Node colours indicate the hierarchical clustering groups.

Users can therefore investigate whether documents assigned to the same hierarchical cluster also tend to appear near one another in the network, whether particular documents act as bridges between groups, and whether highly central documents are associated with particular categories.

```{r fig.height=20, fig.width=20}
# Define k colours for each hierarchical clustering group
cols = c('cadetblue4', 'azure3', "slategray3", "slategrey")
grp_col = cols[groups]

# Plot improved network
plot(
  abs_single, 
  vertex.color = grp_col, 
  layout=layout.fruchterman.reingold, 
  main="Document single node network", 
  # Larger nodes represent greater eigenvector centrality
  vertex.size = eig_abs*15, 
  vertex.label.color = 'black', 
  vertex.label.font=2, 
  vertex.label.cex=1, 
  # Edge width represents the number of shared terms
  edge.width = as.numeric(E(abs_single)$weight)*0.2, 
  edge.color = "grey67", 
  vertex.shape = 'square')

legend('topleft',
       legend = unique(as.factor(groups)), 
       col = unique(grp_col), 
       pch=20, bty = 'n', 
       pt.cex = 5, 
       cex = 2.5, 
       title = 'Hierarchical clustering groups', 
       xjust = -5)
```

---

# Single-Mode Token Network

This section repeats the network analysis, but the nodes now represent **tokens rather than documents**.

In this network:
* nodes = tokens;
* edges = tokens occurring in the same documents;
* edge weight = the number of documents in which the pair of tokens co-occurs.

This allows the analysis to investigate relationships between words rather than relationships between documents.


```{r}
# Multiply transpose binary matrix by binary matrix, this calculates how many documents contain each pair of tokens
tokmatrix = t(dtmsx) %*% dtmsx
# Remove self-connections
diag(tokmatrix) = 0
knitr::kable(tokmatrix)

# Create a weighted, undirected token network
tok_single <- graph_from_adjacency_matrix(tokmatrix, mode = "undirected", weighted = TRUE)

# Network summary
cat('Is simple?',is_simple(tok_single))
cat('\nIs connected?',is_connected(tok_single))
cat('\nDiameter?', diameter(tok_single))
cat('\nAvg path length?',mean_distance(tok_single))
cat('\nEdge density?',edge_density(tok_single))
cat('\nTransitivity?',transitivity(tok_single))

# Centrality measures
bet_tok = betweenness(tok_single)
close_tok = closeness(tok_single)
deg_tok = degree(tok_single)
eig_tok = eigen_centrality(tok_single)$vector
tok_table = cbind(deg_tok, bet_tok, close_tok, eig_tok)
cat('\nSize of largest clique?',clique_num(tok_single))
knitr::kable(tok_table)
```

## Token community detection
Community detection algorithms can be used to identify groups of strongly connected tokens. Two algorithms are demonstrated below. Different algorithms may produce different community structures based on different optimisation strategies and assumptions. The resulting groups should therefore be interpreted in conjunction with the actual token meanings and network structure.

```{r fig.height=20, fig.width=20}
# Edge betweenness community detection
ceb_tok = cluster_edge_betweenness(tok_single)
# Fast greedy community detection
cfg_tok = cluster_fast_greedy(tok_single)

# Plot both community detection methods.
par(mfrow = c(2, 2))
plot(
  ceb_tok,
  tok_single,
  main = "Edge Betweenness")
plot(
  cfg_tok,
  tok_single,
  main = "Fast Greedy")
```

The improved fast greedy community network can be examined for:

* highly central tokens;
* tokens that connect otherwise separate groups;
* strongly co-occurring terms;
* groups of semantically or contextually related tokens;
* differences between community structures.

For example, tokens that frequently occur in the same documents will have stronger connections. However, co-occurrence does not necessarily mean that two words are semantically equivalent. Two terms can be strongly connected because they repeatedly appear in the same subject area or type of document.

```{r fig.height=20, fig.width=20}
# Colours represent Fast Greedy community membership
cols2 = c("thistle4", 'lavenderblush3')
cfg_col = cols2[membership(cfg_tok)]

plot(tok_single, 
     layout=layout.fruchterman.reingold, 
     main="Token single node network",  
     # Node size represents eigenvector centrality
     vertex.size = eig_tok*27, 
     vertex.label.color = 'black', 
     vertex.label.cex=2, 
     # Node colour represents Fast Greedy community
     vertex.color = cfg_col, 
     # Edge width represents co-occurrence weight
     edge.width = E(tok_single)$weight*0.01)

legend('topleft', 
       legend = unique(levels(as.factor(membership(cfg_tok)))), 
       col = unique(cfg_col), 
       pch=20, 
       bty = 'n', 
       pt.cex = 5, 
       cex = 2.5, 
       title = 'CFG communities')
```
---

# Bipartite Network

A bipartite network contains two different types of nodes. For this corpus:
* one type of node = documents;
* second type of node = tokens.

Edges connect documents to the tokens they contain. This allows relationships between the two types of entities to be examined simultaneously. Unlike the document network, where documents are connected to documents, or the token network, where tokens are connected to tokens, the bipartite network directly shows **which tokens occur in which documents**.

```{r}
# Transforming to suitable format
dtms_df = as.data.frame(dtmsx)
dtms_df$ABS = rownames(dtms_df)
dtmsb2 = data.frame()

# Loop through every document
for (i in 1:nrow(dtms_df)){
  for (j in 1:(ncol(dtms_df)-1)){
    touse <- cbind(dtms_df[i,j], dtms_df[i,ncol(dtms_df)], colnames(dtms_df[j]))
  dtmsb2 = rbind(dtmsb2, touse ) 
  } 
}
colnames(dtmsb2) = c("weight", "docs", "token")

# Remove document-token combinations where the token does not occur in the document
dtmsc = dtmsb2[dtmsb2$weight != 0,] 
dtmsc = dtmsc[,c(2,3,1)]

dtmsc
```

### Important consideration for bipartite networks
The bipartite graph below provides direct information about document-token relationships. Documents with many connections have a high degree because they contain many of the retained tokens. Tokens with many connections occur across many documents.

The physical position of nodes in a force-directed layout can help reveal patterns, but proximity should be interpreted together with the actual edges and weights.In this case, centrality values should be interpreted with awareness that the network contains two different node types. For example, comparing the raw degree of a document and a token may not have the same substantive meaning because documents and tokens represent different entities.

The more useful comparison may therefore be:
* which documents have unusually high degree compared with other documents;
* which tokens have unusually high degree compared with other tokens;
* which documents/tokens act as bridges;
* which documents and tokens occupy central positions.

```{r fig.height=15, fig.width=20}
# Create graph from the document-token edge list
g <- graph.data.frame(dtmsc, directed=FALSE)
V(g)$type <- bipartite_mapping(g)$type
V(g)$shape <- ifelse(V(g)$type, "circle", "square")
V(g)$color <- ifelse(V(g)$type, "lightgreen", "pink")
plot(g, layout = layout.fruchterman.reingold)

# Network statistics
cat('\nIs simple?',is_simple(g))
cat('\nIs connected?',is_connected(g))
cat('\nDiameter?', diameter(g))
cat('\nAvg path length?',mean_distance(g))
cat('\nEdge density?',edge_density(g))
cat('\nTransitivity?',transitivity(g))

# Centrality measures
bet_bip = betweenness(g)
close_bip = closeness(g)
deg_bip = degree(g)
eig_bip = eigen_centrality(g)$vector
bip_table = cbind(deg_bip, bet_bip, close_bip, eig_bip)
knitr::kable(bip_table)

```

The improved graph can be interpreted using several visual features simultaneously.

**Node shape**
Different shapes distinguish documents from tokens.

**Node size**
Node size is based on degree, so larger nodes have more connections. For documents, this indicates connections to more retained tokens. For tokens, this indicates occurrence across more documents.

**Edge width**
Thicker edges represent higher document-token frequencies.

**Node colour**
Colour represents Fast Greedy community membership.

**Position**
The force-directed layout attempts to position strongly connected entities closer together. These features allow the user to identify clusters of documents and tokens and investigate which terms characterise particular groups. The actual meaning of each community should be determined by examining the documents and tokens belonging to it.

```{r fig.height=20, fig.width=20}
# Fast Greedy can be used to identify communities in the bipartite network
cfg_bip <- cluster_fast_greedy(g)

# Define colours for communities and assign according to community membership
cols3 = c('lightsteelblue',"honeydew3","seashell3", 'mistyrose4', 'paleturquoise4')
cfg_col2 = cols3[membership(cfg_bip)]

# Improved bipartite graph
g <- graph.data.frame(dtmsc, directed=FALSE)
# Identify node types
V(g)$type <- bipartite_mapping(g)$type
# Colour according to community
V(g)$color <- cfg_col2
# Documents and tokens use different shapes
V(g)$shape <- ifelse(V(g)$type, "circle", "square")
V(g)$label.color = "black"
# Node size represents degree
V(g)$size = deg_bip*0.5
# Different label formatting for the two node types
V(g)$label.font = ifelse(V(g)$type,3,1)
V(g)$label.cex = ifelse(V(g)$type,1.2,1.3)
E(g)$color <- "grey67"
# Edge width represents document-token frequency
E(g)$width = as.numeric(E(g)$weight)*0.5

plot(g, layout = layout.fruchterman.reingold)
legend('topleft', 
       legend = unique(levels(as.factor(membership(cfg_bip)))), 
       col = unique(cfg_col2), 
       pch=20, 
       bty = 'n', 
       pt.cex = 5, 
       cex = 2.5, 
       title = 'CFG communities')
```

---

# Optional improvement: TF-IDF weighting

The current analysis uses term frequencies and shared-term counts. A possible extension is to use **term frequency-inverse document frequency (TF-IDF)** weighting. TF-IDF increases the relative importance of terms that are frequent within particular documents but less common across the entire corpus. This can reduce the influence of very common terms and highlight terms that are more informative for distinguishing documents. TF-IDF can be particularly useful when the objective is to identify terms that distinguish documents rather than simply terms that occur frequently.

```{r}
# Apply TF-IDF weighting to the reduced DTM.
dtm_tfidf <- weightTfIdf(dtms)

# Convert to a standard matrix for inspection.
dtm_tfidf_matrix <- as.matrix(dtm_tfidf)

knitr::kable(dtm_tfidf_matrix)
```

---

# Overall conclusion template

The following structure can be adapted once the results of the user's own corpus have been produced:

The text processing stage transformed the original documents into a format suitable for quantitative text analysis. The preprocessing steps were selected based on the characteristics of the corpus, with unnecessary formatting and common terms removed while attempting to retain terms relevant to the research question.

The resulting Document-Term Matrix provided the basis for the subsequent analyses. Sparse-term removal reduced the number of terms retained in the matrix, allowing the most relevant terms to be examined more clearly.

Hierarchical clustering used cosine distance to measure differences between documents based on their term-use patterns. The resulting dendrogram demonstrated the relationships between documents and allowed groups of similar documents to be identified. Where known categories were available, the resulting clusters were compared with those categories to evaluate how closely the automatically generated groups corresponded to the expected classification.

Sentiment analysis provided a different perspective by examining the emotional or linguistic characteristics of individual documents. Differences between groups were investigated using descriptive plots and statistical tests. The statistical results should be interpreted using both the p-values and the observed differences between groups.

The document network represented shared terms as weighted connections between documents, allowing centrality and connectivity to be investigated. The token network instead examined relationships between terms based on their co-occurrence across documents. Finally, the bipartite network combined documents and tokens into a single network, allowing direct document-token relationships to be examined.

Together, these analyses provide complementary information about the corpus. Clustering is useful for identifying groups of similar documents, while network analysis provides additional information about individual relationships, centrality and community structure. The bipartite network further allows relationships between documents and their constituent terms to be examined directly.

Potential improvements include refining corpus-specific preprocessing, experimenting with different sparsity thresholds, comparing alternative clustering methods, investigating different sentiment dictionaries and applying TF-IDF weighting. These choices should be guided by the research question and characteristics of the corpus rather than applied automatically.
